Variants on the Flight and Maintenance Planning problem

### 2019-04-09 latex/notes/model/model.tex
renamed some sets and params according to Robert's comments. Corrected some minor errors.

        \begin{tabular}{p{5mm}p{120mm}}
            $\mathcal{I}$     &  aircraft. \\
            $\mathcal{T}$     &  time periods included in the planning horizon. \\
            $\mathcal{J}$     &  missions. \\
            $\mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
        \end{tabular}

@@

        \begin{tabular}{p{10mm}p{115mm}}
            $H_j$             & amount of flight time required by mission $j$. \\
            $R_j$             & number of aircraft required by mission $j$. \\
            $MT_j$            & minimum number of consecutive periods an aircraft has to be assigned to mission $j$. \\
            $U^{min}$       & default aircraft flight time if it is not assigned to any mission nor in maintenance.\\
            $M$               & check duration in number of periods. \\
        \end{tabular}

@@

        \begin{tabular}{p{10mm}p{115mm}}
            $C^{max}$         & maximum number of simultaneous checks. \\
            $E^M$         & maximum number of periods between two consecutive checks. \\
            $E^m$         & minimum number of periods between two consecutive checks. \\
            $H^M$               & remaining flight time after a check. \\
        \end{tabular}

@@

        \begin{tabular}{p{10mm}p{115mm}}
            $N_t$             & number of aircraft known to be in maintenance in period $t$.\\
            $N^K_{kt}$         & number of aircraft in cluster $k$ known to be in maintenance in period $t$.\\
            $A^K_{kt}$         & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$.\\
            $H^K_{kt}$         & minimum number of total remaining flight time for cluster $k$ at period $t$.\\
            $Rft^{Init}_i$  & remaining flight time for aircraft $i$ at the start of the planning horizon. \\
            $Rct^{Init}_i$  & remaining calendar time for aircraft $i$ at the start of the planning horizon. \\
        \end{tabular}

@@

        \begin{tabular}{p{5mm}p{120mm}}
            $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
            $\mathcal{J}_t $    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
            $\mathcal{I}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
            $\mathcal{I}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
            $\mathcal{O}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
            $\mathcal{A}^{init}_j$  & aircraft $i \in \mathcal{I}$ that have mission $j$ pre-assigned in the previous period to the start of the planning horizon. \\
        \end{tabular}

@@

        \begin{tabular}{p{8mm}p{117mm}}
            $\mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
            % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $\mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^m-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $\mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^M - M$ such that $t' \in \{ t + M + E^m-1 , ...,  t + M + E^M-1 \}$ (figure \ref{fig:gantt_windows}a). \\
            $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^M + E^m \}}\}$ (figure \ref{fig:gantt_windows}b). \\
            $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^M + E^m \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
            $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
            $\mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{T}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
            $\mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
        \end{tabular}

@@

    \begin{align}
        & \text{Min}\; \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}} m_{it_1t_2}  \label{eq:objective1}\\
        & \text{Min}\; \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}} m_{it_1t_2} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
    \end{align}

@@

    The following constraints are used in the model:
    \begin{align}
        % maximum capacity1:
        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} \sum_{i \in \mathcal{I}} m_{it_1t_2} + N_t \leq C^{max}
          & t \in \mathcal{T} \label{eq:capacity1}\\
        % min assignments:
        & \sum_{i \in \mathcal{I}_j} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:missionres}\\
        % just doing one thing at any given time:
        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

@@

%     \begin{align}
%         % if start mission => assign start
%         & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
%                 & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
%         & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_j}
%                 & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\
%         & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
%         & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
%     \end{align}

@@

    \begin{align}
       & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} \sum_{i \in \mathcal{I}_k} m_{it_1t_2} + N^K_{kt} \leq A^K_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
       & \sum_{i \in \mathcal{I}_k} rft_{it} \geq H^K_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
    \end{align}

@@

    \begin{align}
        % remaining used time
         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
         & u_{it} \in [U^{min}, \max_j{\{H_j\}}]
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
         & rft_{it} \leq rft_{i(t-1)} + H^M \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
        & rft_{i0} = Rft^{Init}_i
               & i \in \mathcal{I} \label{eq:rft_initial}\\
        & rft_{it} \geq H^M \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
                &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
        & rft_{it} \in [0,H^M]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
    \end{align}

@@

    \begin{align}
       % For now, we only want one or possibly two maintenances...
        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
          & i \in \mathcal{I}\label{eq:num_maint} \\
        % & m_{it'} + m_{it} \leq 1
        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\ 
        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
        % & m_{it_1t_2} = 0
        %   & t_1 \in \mathcal{T}^{m_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
        & m_{it_1t_2} =  0
          & t_1 \notin \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
        & m_{it_1t_2} =  0
          & t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I} \label{eq:maint_forbid2}
    \end{align}

>>>>

        \begin{tabular}{p{15mm}p{110mm}}
            $i \in \mathcal{I}$     &  aircraft. \\
            $t \in \mathcal{T}$     &  time periods included in the planning horizon. \\
            $j \in \mathcal{J}$     &  missions. \\
            $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
        \end{tabular}

@@

        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
            $H_j$             & amount of flight time required by mission $j$. & hours. \\
            $R_j$             & number of aircraft required by mission $j$. & aircraft. \\
            $MT_j$            & minimum number of consecutive periods an aircraft has to be assigned to mission $j$. & periods. \\
            $U^{min}$         & default aircraft flight time if it is not assigned to any mission nor in maintenance.& hours \\
            $M$               & check duration in number of periods. & periods. \\
        \end{tabular}

@@

        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
            $C^{max}$         & maximum number of simultaneous checks. & aircraft. \\
            $E^{max}$         & maximum number of periods between two consecutive checks. & periods. \\
            $E^{min}$         & minimum number of periods between two consecutive checks. & periods. \\
            $H^{max}$         & remaining flight time after a check. & hours. \\
        \end{tabular}

@@

        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
            $N_t$               & number of aircraft known to be in maintenance in period $t$. & aircraft. \\
            $N^{Clust}_{kt}$    & number of aircraft in cluster $k$ known to be in maintenance in period $t$. & aircraft. \\
            $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
            $H^{Clust}_{kt}$    & minimum number of total remaining flight time for cluster $k$ at period $t$. & hours. \\
            $Rft^{Init}_i$      & remaining flight time for aircraft $i$ at the start of the planning horizon.  & hours.  \\
            $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
        \end{tabular}

@@

        \begin{tabular}{p{15mm}p{110mm}}
            $t \in \mathcal{TJ}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
            $j \in \mathcal{JT}_t $    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
            $i \in \mathcal{IJ}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
            $i \in \mathcal{IK}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
            $j \in \mathcal{JI}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
            $i \in \mathcal{A}^{Init}_j$  & aircraft $i \in \mathcal{I}$ that have mission $j$ pre-assigned in the previous period to the start of the planning horizon. \\
        \end{tabular}

@@

        \begin{tabular}{p{30mm}p{95mm}}
            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
            % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
        \end{tabular}

@@

    \begin{align}
        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|} m_{it_1t_2}  \label{eq:objective1}\\
        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1}} m_{it_1t_2} \times H^{max} - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
    \end{align}

@@

    The following constraints are used in the model:
    \begin{align}
        % maximum capacity1:
        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + N_t \leq C^{max}
          & t \in \mathcal{T} \label{eq:capacity1}\\
        % min assignments:
        & \sum_{i \in \mathcal{IJ}_j, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
        % just doing one thing at any given time:
        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

@@

%     \begin{align}
%         % if start mission => assign start
%         & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
%                 & t =1, ..., \mathcal{T}, j \in \mathcal{JT}_t, i \in \mathcal{IJ}_j \label{eq:start1} \\
%         & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{Init}_j}
%                 & j \in \mathcal{J}_0, i \in \mathcal{IJ}_j \label{eq:start2} \\
%         & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
%         & j \in \mathcal{J}, t \in \mathcal{TJ}_j, i \in \mathcal{IJ}_j \label{eq:start3}
%     \end{align}

@@

    \begin{align}
       & \sum_{i \in \mathcal{IK}_k, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
       & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
    \end{align}

@@

    \begin{align}
        % remaining used time
         & u_{it} \geq \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
         & u_{it} \in [U^{min}, \max_j{\{H_j\}}]
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
         & rft_{it} \leq rft_{i(t-1)} + H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
        & rft_{i0} = Rft^{Init}_i
               & i \in \mathcal{I} \label{eq:rft_initial}\\
        & rft_{it} \geq H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
                &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
        & rft_{it} \in [0,H^{max}]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
    \end{align}

@@

    \begin{align}
       % For now, we only want one or possibly two maintenances...
        & \sum_{t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}| } m_{it_1t_2} =  1 
          & i \in \mathcal{I}\label{eq:num_maint} \\
        % & m_{it'} + m_{it} \leq 1
        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\ 
        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
        % & m_{it_1t_2} = 0
        %   & t_1 \in \mathcal{T}^{m_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
        & m_{it_1t_2} =  0
          & t_1 \notin \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
        & m_{it_1t_2} =  0
          & t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup |\mathcal{T}|, i \in \mathcal{I} \label{eq:maint_forbid2}
    \end{align}
